---
title: "01_data_profiling_complete"
author: "Crystal"
output: 
  rmarkdown::html_document:
    toc: true
    toc_float: true
    highlight: tango
    theme: journal
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, include=FALSE, paged.print=FALSE}
# load packages 
for (pkg in c("tidyverse", "data.table", "stringr", "stringi", "mosaic", "dplyr", "readr","gt")) {
  library(pkg, character.only = TRUE)
}

get_db_conn <-
  function(db_name = "sdad",
           db_host = "postgis1",
           db_port = "5432",
           db_user = Sys.getenv("db_usr"),
           db_pass = Sys.getenv("db_pwd")) {
    RPostgreSQL::dbConnect(
      drv = RPostgreSQL::PostgreSQL(),
      dbname = db_name,
      host = db_host,
      port = db_port,
      user = db_user,
      password = db_pass
    )
  }

con <- get_db_conn()


pers <- DBI::dbGetQuery(con, "
                        SELECT * FROM bgt_res.pers
    "
)

DBI::dbDisconnect(con)

uszips <- read_csv("~/git/dspg20career/src/burningglass/uszips.csv")%>%
  select(zip, city, state_id, state_name, county_name)

```

# Objective
Provide data profiling on the complete burning glass dataset. 

# Summary
In pers table, `r sum(is.na(pers$msa))/nrow(pers)`% (N= `r sum(is.na(pers$msa))`) missing in the msa variable, which is an identifier for geographic location. `r length(grep("47900", pers$msa, fixed = T))`users came from Washington-Arlington-Alexandria, DC-VA-MD-WV Metropolitan Statistical Area, identified by code: 47900. 

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
# function for data profiling
data_profiling <- function(df){
summary_table <- tibble(var = names(df),
       variable_type = map(.x = df, .f = function(col) class(x = col)),                 
       num_unique = map_int(.x = df, .f = function(col) length(x = unique(x = col))),
       num_missing = map_int(.x = df, .f = function(col) sum(x = is.na(x = col)))) %>%
  mutate(perc_missing = round(x = 100 * num_missing / nrow(x = .), digits = 2L))

return(summary_table)
}
```

```{r message=FALSE, warning=FALSE}
data_profiling(pers)
```

# Identify DC Metropolitan Area

## MSA
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
#library(zipcode)

DC_pers <- pers%>%
  filter(grepl("47900", msa) == T)

matches <- regmatches(DC_pers$msa, gregexpr("[[:digit:]]+", DC_pers$msa))
msa_list <- unique(unlist(matches))
msa_DC <- c("47900", "548")
msa_not_DC <- setdiff(msa_list, msa_DC)

dup_detect <- c()
for (i in (1:length(matches))){
  item <- matches[[i]]
  if(sum(item %in% msa_not_DC == T) > 2){
    dup_detect[i] = T
  }else{
    dup_detect[i] = F
  }
}

```

```{r}
library(maditr)
msa <- fread(input = "https://www2.census.gov/programs-surveys/cbp/technical-documentation/reference/metro-area-geography-reference/msa_county_reference17.txt")
# https://www2.census.gov/programs-surveys/cbp/technical-documentation/reference/state-county-geography-reference/georef17.txt for States + Counties (FIPS)
```

We identified 687129 people who haved listed "47900" in their msa, 0.30 % (N=2093) of people had duplicated entries in msa. Among people who only reported "47900" in their msa, there are 5033 unique self-reported zipcodes. Then, I matched the zipcode with zipcodes on US Zip Codes Database (using authoritative sources including the U.S. Postal Service™, U.S. Census Bureau, National Weather Service, American Community Survey, and the IRS). 28.00% of these 5033 zipcodes (N=1409) don't match with the zipcode database. 
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
nrow(DC_pers)

table(dup_detect) #2093 ppl had msa duplicates

DC_pers <- cbind(DC_pers, dup_detect)

DC_pers <- DC_pers%>%
  filter(dup_detect == F)%>%
  mutate(zipcode_clean = substr(zipcode,1,5))

DCzipcode <- DC_pers%>%
  group_by(zipcode_clean)%>%
  summarize(N=n())%>%
  left_join(uszips, by = c("zipcode_clean" = "zip"))%>%
  arrange(desc(N))

table(is.na(DCzipcode$state_id))
nrow(DCzipcode)

DCzipcode%>%
  group_by(state_id)%>%
  summarize(number_zips=n())%>%
  arrange(desc(number_zips))%>%
  gt()%>%
   tab_header(
    title = "states zipcode in 47900 msa code"
  )


```

# Subset Burning Glass data using DC/VA/MD zipcodes
Alternatively, we could us zipcodes identified in the zipcode database. There are 52 zipcodes identified in the District of Columbia. We could add more states if we want. For example 893 zipcodes from VA, 468 zipcodes from Maryland. In total, there are 1416 zipcodes in DC/VA/MD. In Burning Glass data, 1,307,546 people are located in DC/VA/MD, 84,337 people located in DC using zipcode identifier.
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
table(uszips$state_id)

# only DC
dc_zips <- uszips%>%
  filter(state_id == "DC")
dc_zips <- as.vector(dc_zips$zip)

pers_dc <- pers%>%
  filter(zipcode %in% dc_zips)
nrow(pers_dc)

# DC, VA, MD
dc_va_md_zips <- uszips%>%
  filter(state_id %in% c("DC", "VA", "MD"))

dc_va_md_zips <- as.vector(dc_va_md_zips$zip)
length(dc_va_md_zips)

pers_dc_va_md <- pers%>%
  filter(zipcode %in% dc_va_md_zips)
nrow(pers_dc_va_md)
```


