---
title: "01_data_profiling"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
# load packages 
for (pkg in c("tidyverse", "igraph", "visNetwork", "data.table", "R.utils", "RPostgreSQL", "cowplot", "maditr", "stringr", "stringi", "mosaic", "dplyr", "ggplot2")) {
  library(pkg, character.only = TRUE)
}

get_db_conn <-
  function(db_name = "sdad",
           db_host = "postgis1",
           db_port = "5432",
           db_user = Sys.getenv("db_usr"),
           db_pass = Sys.getenv("db_pwd")) {
    RPostgreSQL::dbConnect(
      drv = RPostgreSQL::PostgreSQL(),
      dbname = db_name,
      host = db_host,
      port = db_port,
      user = db_user,
      password = db_pass
    )
  }

con <- get_db_conn()

cert <- DBI::dbGetQuery(con, "
                        SELECT * FROM bgt_res.cert
                        WHERE cert.id in (
                        SELECT ID FROM bgt_res.pers
                        WHERE pers.msa like '%47900%'
                        )"
)

ed <- DBI::dbGetQuery(con, "
                      SELECT * FROM bgt_res.ed
                      WHERE ed.id in (
                      SELECT ID FROM bgt_res.pers
                      WHERE pers.msa like '%47900%'
                      )"
)

DBI::dbDisconnect(con)
```


```{r}
# function for data profiling
data_profiling <- function(df){
  variable_type <- apply(df, 2, function(x) class(x))
  num_unique <- apply(df, 2,  function (x) length(unique(x)))
  num_missing <- apply(df, 2, function(x) sum(is.na(x)))
  perc_missing <- apply(df, 2, function(x) round((sum(is.na(x)))/length(x)* 100, digits = 2))  
  summary_table <- cbind(variable_type, num_unique, num_missing, perc_missing)
  return(summary_table)
}
```


# Table: Cert
In total, there are `r nrow(cert)` entries in cert table. There are two types of certificates: certification and license. On average, each person received 2 certificates (sd = 1.68). 
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
data_profiling(cert)

cert <- cert%>%
  mutate(type = if_else(type == "license", "License", type))

table(cert$type)

cert_n <- cert%>%
  group_by(id)%>%
  summarize(n_cert = n())

ggplot(cert_n, aes(x = n_cert)) + 
  geom_histogram(binwidth = 1.3) + 
  labs(title = "Distribution of the number of certificates people received", x = "number of certificates") 
favstats(cert_n$n_cert)

# top 10 certificates
cert_name <- cert%>%
  group_by(name)%>%
  summarize(n_cert_name = n())%>%
  arrange(desc(n_cert_name))%>%
  top_n(10)%>%
  rename("number of people who received the certificate" = "n_cert_name")
```



# Table: Ed
In total, there are `r nrow(ed)` entries in ed table. After a preliminary clean up of degree, we identified `r round((nrow(degree_clean)-267101)/(nrow(degree_clean))* 100, digits = 2) `% of self reported degree type into 5 categories: some high school, high school, bachelor, master, and doctor's degree. 
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
data_profiling(ed)
table(ed$degreetype)

degree_clean <- ed%>%
  filter(!is.na(degreetype))%>%
  mutate(degree_somehs = if_else(str_detect(string = degreetype, 
                              pattern = "\\b(?i)(10|11|9)\\b"), T,F))%>%
  mutate(degree_hs = if_else(str_detect(string = degreetype, 
                              pattern = "\\b(?i)(12|High School|ged)\\b"), T,F))%>%
  mutate(degree_bachelor = if_else(str_detect(string = degreetype, 
                              pattern = "\\b(?i)(Bachelor|bachelor|Bachelors|BS|bs|AA|A.A|Undergraduate|undergraduate|postgraduate|Associate|associate)\\b"), T,F))%>%
  mutate(degree_master = if_else(str_detect(string = degreetype, 
                              pattern = "\\b(?i)(master|Master|MBA|M.S|MS|MD)\\b"), T,F))%>%
  mutate(degree_doctor = if_else(str_detect(string = degreetype, 
                              pattern = "\\b(?i)(phd|Ph.D|postdoc)\\b"), T,F))%>%
  mutate(degree_highest = if_else(degree_doctor == T, "doctor", 
                                  if_else(degree_master == T, "master",
                                          if_else(degree_bachelor == T, "bachelor", if_else(degree_hs == T, "hs", 
                                                                                            if_else(degree_somehs == T, "somehs", "others"))))))



table(degree_clean$degree_highest)
```

